A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-07-31 tex/paper.tex
skip runs that haven't finished

\subsection{Generic Tests}

In addition to running tests under the same assumptions as historical analyses, we also allow the region kernels to play a role.

Note that using global covariance actually lets the parameters float a bit more, and provides some cushion against the tyranny of too many lines. Only so many parameters, can't infinitely tweak the model spectrum for any large region of spectrum. This parameterizes the badness of fit, better models will have a lower covariance, as evidenced by the Kurucz (-14.6) v. PHOENIX (-14.2) global cov amplitudes.

@@

@@

@@

@@

\begin{itemize}
  \item Spectral parameters can vary by a large margin depending on which spectral library you use ($200~K$ or more). And which technique, which may be a result of choosing certain indices, certain models, or both. Therefore it is in our interest to use as wide a region of spectrum as possible.
  \item Both spectral libraries have stars that they perform better and worse on.
  \item This will be reflected in the increased level of global noise, and number of ``bad'' regions that have been instantiated.
\end{itemize}

Summary end of this section is that we are

@@

@@

\begin{deluxetable}{llrlr@{$\pm$}lr@{$\pm$}lr@{$\pm$}ll} 
\tablecaption{\label{table:tests} Results}
\tablehead{
  \colhead{Object} & \colhead{Orders} & \colhead{$\lambda$ range [\AA]} & \colhead{library} &
  \multicolumn{2}{c}{$T_{\rm eff} \pm \sigma$} & \multicolumn{2}{c}{$\log g \pm \sigma$} &
  \multicolumn{2}{c}{$\Z \pm \sigma$} & \colhead{comments}
}
\startdata
WASP-14 & 23 & & Kurucz & 6241 & 15 & 3.88 & 0.02 & -0.43 & 0.01 & Poisson\\
WASP-14 & 23 & & Kurucz & 6264 & 31 & 3.92 & 0.04 & -0.42 & 0.02 & Global\\
WASP-14 & 23 & & Kurucz & & & \multicolumn{2}{c}{\nodata} & & & Matern; fixed $\log g = 4.29$\\
WASP-14 & 23 & & PHOENIX & 6023.3 & 4.3 & 3.84 & 0.01 & -0.50 & 7.7e-4 & Poisson\\
WASP-14 & 23 & & PHOENIX & 6000.72 & 3 & 3.94 & 0.01 & -0.50 & 2.6e-3 & Poisson, masked \\
WASP-14 & 23 & & PHOENIX & 5984.9 & 74 & 3.80 & 0.09 & -0.66 & 0.07 & Global, no mask \\
WASP-14 & 23 & & PHOENIX & & & & & & & Global and local \\
WASP-14 & 22-24 & 5060-5315 & Kurucz & & & & & & & Poisson\\
WASP-14 & 22-24 & 5060-5315& Kurucz & & & & & & & Global\\
WASP-14 & 22-24 & 5060-5315& Kurucz & & & \multicolumn{2}{c}{\nodata} & & & \tablenotemark{a}Global; fixed $\log g = 4.29$\\
WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Poisson\\
WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global; low because of 24\\
WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Global; fixed $\log g = 4.29$ \\
WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global and local\\
\enddata
\tablenotetext{a}{Using the same assumptions in the \citet{torres12} {\tt SPC} analysis: same region of spectrum, same spectral library, and fixing $\log g$=$4.29$, the value determined by \citet{joshi09} from the exoplanet transit. For the WASP-14 spectrum acquired by \emph{TRES}, \citet{torres12} use {\tt SPC} to determine $T_{\rm eff}$=6507$\pm$50K and $\Z$=-0.15$\pm$0.08. With {\tt MOOG}, they determine $T_{\rm eff}$=6150$\pm$75K and $\Z$=-0.34$\pm$0.10. Because {\tt SPC} uses a cross-correlation technique which only accounts for the white noise and we model residual spectrum as correlated noise, there may be differences in the uncertainty estimate.}
\end{deluxetable}

>>>>

\subsection{WASP-14} \label{subsec:wasp}
To evaluate the performance of our methodology, we perform a series of tests using WASP-14 and Gl~51. We use the PHOENIX grid (made with version XX PHOENIX, \citealt{husser13}), which spans wavelengths from XX to XX. The grid is very high quality and has been hand-calibrated against the Sun and Vega, therefore it performs very well for solar like stars. Because WASP-14 is an optical spectrum, we can use the high-quality CfA Kurucz grid, synthesized by Laird and Morse and used by {\tt SPC}. Because the spectrum of Gl~51 is the mid-infrared, we are unable to use the CfA grid for these comparisons.

@@

For demonstrative purposes, the first fits we perform is using only the Poisson variances from the original spectrum (WK1 and WP1), ignoring any correlations that might arise in the residual spectrum. This is equivalent to setting $b=1$ and all other amplitudes of $\cov$ to 0, and allows our likelihood to simplify to the familiar form of $\chi^2$. While the uncertainty estimate returned from these fits is remarkable, it also leads to the situation of overconfidence. Such a small uncertainty arises because such a likelihood function assumes that the data set is generated from the model, meaning that the only way in which the data set might differ from the model is through the addition of white Poisson noise. Unfortunately, this is incorrect. While the synthetic stellar model is \emph{incredibly good}, it is not a perfect representation of the star, nor could it be given that a synthetic spectrum from a grid has only 3 $\vt_\ast$ to tweak. This means that some line strengths might be different, whether due to slight inaccuracies in the atomic constants or metallicity compositions that differ in a few elements different from what the Sun might be. Thus, the data set differs from the model not just from added Poisson uncertainties, but also \emph{systematically} because the model cannot be a perfect representation of reality. This is equivalent to fitting a straight line to a series of data points whose variance is drastically underestimated. A linear least squares fit $\chi^2$ routine might return a small uncertainty estimate, however the slope and intercept of the line might be drastically biased. Therefore, it is clear that they model requires a way to account for this additional source of ``noise.'' Make the distinction that the Kurucz model fits \emph{very} well, but still requires some global covariance simply because of the three parameter problem.

Spectral parameters can vary by a large margin depending on which spectral library you use ($200~K$ or more). And which technique, which may be a result of choosing certain indices, certain models, or both. Therefore it is in our interest to use as wide a region of spectrum as possible. Both spectral libraries have stars that they perform better and worse on. This will be reflected in the increased level of global noise, and number of ``bad'' regions that have been instantiated. Eg, WASP 14 v. Gl~51.

@@

Because the PHOENIX spectrum is computed over a wide range, there are often some egregious ``outlier'' lines that appear in the spectrum. To compare our approach with the common approach of ``sigma-clipping,'' for the next test (WP2), we manually mask any region of spectrum which has a more than 3 sigma residual near the best fit parameters. As you can see, this has the effect of actually changing the final stellar estimate significantly, while also reducing the uncertainty in the stellar parameters. Sigma-clipping like this is a procedural approach and is not justified in a probabilistic framework \citep{hogg10}.

For the next tests, (WK2 and WP3), we introduce the global covariance kernel to the fit. This allow the model-data mismatch, and we can see that the uncertainty estimate naturally inflates to capture the discrepancy. 
Note that using global covariance actually lets the parameters float a bit more, and provides some cushion against the tyranny of too many lines. Only so many parameters, can't infinitely tweak the model spectrum for any large region of spectrum. This parameterizes the badness of fit, better models will have a lower covariance, as evidenced by the Kurucz (-14.6) v. PHOENIX (-14.2) global cov amplitudes.
I would like to point out that this came about \emph{self-consistently} through the model, not something that was added in post-facto as is commonly done (cough, \citep{torres12, buchhave12}). Because our uncertainty is derived through the model itself, it \emph{preserves correlations} between parameter estimates. Here we see that $\vt_\ast$ are highly correlated. If you just slapped on 50K errorbars to everything, you would be destroying this information.

For the next tests, (), we introduce the local covariance kernels to track outlier lines. These kernels can map out the regions of the spectrum which are outliers. The glory of these local kernels are that the parameters allow to play a role in data driven spectra. See figure XX for a cool visualization of what it looks like.

@@

\subsection{Gl~51}
Gl~51 tests. Using no covariance, covariance, just the lines, just same region of spectrum. Adding local covariance kernels to Na and Ca lines that are underpredicted.

@@

@@

Adding in additional orders to the spectrum allows the posterior to broaden (if each order has disparate predictions about the stellar parameters) or to contract, since we are also fitting for the model-data mismatch. If all orders match well (case of Kurucz), then the global kernels will have low amplitude. However, if certain orders predict different stellar parameters, then for the one set of stellar parameters that is shared by the dataset, the global kernels will inflate for the other orders to absorb the model-data mismatch.

@@

\begin{deluxetable}{lllrlr@{$\pm$}lr@{$\pm$}lr@{$\pm$}ll} 
\tablecaption{\label{table:tests} Results}
\tablehead{
  \colhead{Test ID} & \colhead{Object} & \colhead{Orders} & \colhead{$\lambda$ range [\AA]} & \colhead{library} &
  \multicolumn{2}{c}{$T_{\rm eff} \pm \sigma$} & \multicolumn{2}{c}{$\log g \pm \sigma$} &
  \multicolumn{2}{c}{$\Z \pm \sigma$} & \colhead{comments}
}
\startdata
WK1 & WASP-14 & 23 & 5135-5235 & Kurucz & 6241 & 15 & 3.88 & 0.02 & -0.43 & 0.01 & Poisson\\
WK2 & WASP-14 & 23 & 5135-5235 & Kurucz & 6264 & 31 & 3.92 & 0.04 & -0.42 & 0.02 & Global\\
WK3 & WASP-14 & 23 & 5135-5235 & Kurucz & & & \multicolumn{2}{c}{\nodata} & & & Matern; fixed $\log g = 4.29$\\
WP1 & WASP-14 & 23 & 5135-5235 & PHOENIX & 6023.3 & 4.3 & 3.84 & 0.01 & -0.50 & 7.7e-4 & Poisson\\
WP2 & WASP-14 & 23 & 5135-5235 & PHOENIX & 6000.72 & 3 & 3.94 & 0.01 & -0.50 & 2.6e-3 & Poisson, masked \\
WP3 & WASP-14 & 23 & 5135-5235 & PHOENIX & 5984.9 & 74 & 3.80 & 0.09 & -0.66 & 0.07 & Global, no mask \\
WP4 & WASP-14 & 23 & 5135-5235 & PHOENIX & & & & & & & Global and local \\
&WASP-14 & 22-24 & 5060-5315 & Kurucz & & & & & & & Poisson\\
&WASP-14 & 22-24 & 5060-5315& Kurucz & & & & & & & Global\\
&WASP-14 & 22-24 & 5060-5315& Kurucz & & & \multicolumn{2}{c}{\nodata} & & & \tablenotemark{a}Global; fixed $\log g = 4.29$\\
&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Poisson\\
&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global; low because of 24\\
&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Global; fixed $\log g = 4.29$ \\
&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global and local\\
GP1 &Gl~51 & \nodata & IR & PHOENIX & & & & & & & Poisson\\
\enddata
\tablenotetext{a}{Using the same assumptions in the \citet{torres12} {\tt SPC} analysis: same region of spectrum, same spectral library, and fixing $\log g$=$4.29$, the value determined by \citet{joshi09} from the exoplanet transit. For the WASP-14 spectrum acquired by \emph{TRES}, \citet{torres12} use {\tt SPC} to determine $T_{\rm eff}$=6507$\pm$50K and $\Z$=-0.15$\pm$0.08. With {\tt MOOG}, they determine $T_{\rm eff}$=6150$\pm$75K and $\Z$=-0.34$\pm$0.10. Because {\tt SPC} uses a cross-correlation technique which only accounts for the white noise and we model residual spectrum as correlated noise, there may be differences in the uncertainty estimate.}
\end{deluxetable}
